\sysfull

### 2017-08-10 paper/coalescing.tex
Covering Koen's comments

\sys can coalesce an arbitrary number consecutive tasks. As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance. As more tasks coalesce, the amount of potentially {\em wasted work} also increases. If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of
the tasks in the long sequence. \sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, adaptively determining a coalescing factor that combines many tasks without suffering an excess of wasted work.

@@

\nonindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks. Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition. Figure~\ref{fig:coalescing_example}(b) shows the same sequence of tasks executing as coalesced groups of three tasks. Figure~\ref{fig:coalescing_example}(c) shows how \sys adapts the value of $C$ based on $S$. \sys executes 12 tasks in the first operating period, assigning $S$ to six. \sys completes only five of the six coalesced tasks in the next operating period and then power fails. \sys sets $S$ to two in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails. Execution resumes at the start of the third pair of tasks with a coalescing factor of two. If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.

>>>>

\sys can coalesce an arbitrary number consecutive tasks. As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance. As more tasks coalesce, the amount of potentially {\em wasted work} also increases. If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of
the tasks in the long sequence. \sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.

@@

\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks. Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition. Figure~\ref{fig:coalescing_example}(b) shows the same sequence of tasks executing as coalesced groups of three tasks. Figure~\ref{fig:coalescing_example}(c) shows how \sys adapts the value of $C$ based on $S$. \sys executes 12 tasks in the first operating period, assigning $S$ to six. \sys completes only five of the six coalesced tasks in the next operating period and then power fails. \sys sets $S$ to two in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails. Execution resumes at the start of the third pair of tasks with a coalescing factor of two. If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.

### 2017-08-10 paper/evaluation.tex
Compiler section revised
Covering Koen's comments

We also had alternative setup where we used MSP-TS430RGZ48C launchpad for continuous power experiment. We used internal timer for measuring run time. For harvested energy, we used ThingMagic Astra-EX RFID reader with power level of 17.25-30dBm. We used EDB~\cite{edb} for run time measurement. Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.

@@

\subsubsection{Task Coalescing Performance}
\label{sec:results_coalescing}

@@

The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e. from energy harvesting to continuous power environment) is the best one?. To answer this question we have implemented a simple program composed of $x$ empty tasks. We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details). The result is presented in Figure~\ref{fig:virtualTaskSize}. We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement). However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time. \todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}

>>>>

We also had alternative setup where we used MSP-TS430RGZ48C launchpad for continuous power experiment. We used internal timer for measuring run time. For harvested energy, we used ThingMagic Astra-EX RFID reader with power level of 17.25-30dBm. We used EDB~\cite{edb} for run time measurement. Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}

@@

%\subsubsection{Task Coalescing Performance}
%\label{sec:results_coalescing}

\vspace{2cm}

List of figures: 
\begin{itemize}
	\item Fig. 1: Viper only/per app/fixed power/(hand/complier)/no coalescing/clang [Resuls are ready by Kiwan]
	\item Fig. 2: Viper only/per app/per distance (fixed/20/40/60)/(hand/complier)/coalescing [Resuls are ready by Kiwan]
	\item Fig. 3a: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/clang [Resuls are ready by Kiwan]
	\item Fig. 3b: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/gcc [Resuls are not ready: Kiwan needs to make them]
	\item Fig. 4: exactly as Fig 12 in the paper (4 apps only)
	\item Fig. 5: average virtual task size/gcc/fixed distance/per app 
\end{itemize}

\vspace{2cm}

@@

The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e. from energy harvesting to continuous power environment) is the best one? To answer this question we have implemented a simple program composed of $x$ empty tasks. We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details). The result is presented in Figure~\ref{fig:virtualTaskSize}. We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement). However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time. \todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}

### 2017-08-10 paper/paging.tex
Covering Koen's comments

\sys uses a page-based virtual memory scheme that allows tasks to primarily use SRAM instead of frequently accessing FRAM. The paging scheme partitions the non-volatile memory address space into pages of {\em protected data}. As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} that is stored in volatile memory and holds a single page. When a modified page is swapped out, it is buffered in a dedicated commit buffer for later commit. \sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures. \sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
hardware DMA block-copy operations that have higher bandwidth and avoid fixed, 
per-variable data movement costs. \sys uses efficient DMA block copies when paging data in and out, as well as during two-phase commit. \sys is unlike prior work~\cite{chain,alpaca} that moves one variable at a time, prohibiting efficient, DMA bulk transfers. Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}. Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.

@@

\sys virtualizes the protected, non-volatile data that a task can access through a modified memory interface. The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}. {\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer \texttt{pageBuf} in volatile memory.  In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}. The page tag identifies in which page of non-volatile memory the memory location resides and
the page offset identifies a particular byte of that page. How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory. To address {\tt p} pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises $log(s)$ bits and $log(p) + log(s) = a$. For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a
page's tag is the high-order nine bits of the address and a byte's page offset
is the remaining 7 bits.

@@

After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer. \sys keeps track of the page tag for the page currently resident in the paging buffer in {\tt CrntPagTag}, which is stored in non-volatile memory. When a task accesses a variable, it compares the variable's page tag to {\tt CrntPagTag}.  If an accessed variable's page is equal to {\tt CrntPagTag} then the accessed variable's page is resident in the page buffer. If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer. If the accessed variable's page is resident in the page buffer, an executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing). If the accessed variable's page is not resident in the page buffer, the operation incurs a {\em page fault}. At a page fault, \sys must swap out the
page that is resident int he page buffer and swap in the the variable's page.

@@

Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism. The task manipulates protected variables {\tt x}, {\tt y}, and {\tt z}, which are in the same page. On accessing {\tt x}, its page is swapped into the page buffer. On accessing {\tt y} and {\tt z}, the task need only access the volatile page buffer. The task updates the page, leaving it dirty at the task's completion. When the task completes, it first copies the page from the page buffer into the non-volatile {\tt
pagesTemp} region before commit. Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.

>>>>

\sys paging scheme partitions the non-volatile memory address space into pages of {\em protected data}. As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} that is stored in volatile memory and holds a single page. When a modified page is swapped out, it is buffered in a dedicated commit buffer for later. \sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures. \sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
hardware DMA block-copy operations, unlike prior work~\cite{chain,alpaca} that moves one variable at a time. Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}. Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.

@@

\sys virtualizes the protected, non-volatile data that a task can access through a modified memory interface. The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}. {\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer \texttt{pageBuf} in volatile memory.  In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}. The page tag identifies in which page of non-volatile memory the memory location resides and the page offset identifies a particular byte of that page.

%How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory. To address {\tt p} pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises $log(s)$ bits and $log(p) + log(s) = a$. For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a page's tag is the high-order nine bits of the address and a byte's page offset is the remaining 7 bits.

@@

After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer. \sys keeps track of the page tag for the page currently resident in the paging buffer in {\tt CrntPagTag}, which is stored in non-volatile memory. When a task accesses a variable, it compares the variable's page tag to {\tt CrntPagTag}.  If an accessed variable's page is equal to {\tt CrntPagTag} then the accessed variable's page is resident in the page buffer. The executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing). If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer; the operation incurs a {\em page fault}. At a page fault, \sys must swap out the page that is resident in the page buffer and swap in the the variable's page.

@@

Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism. The task manipulates protected variables {\tt x}, {\tt y}, and {\tt z}, which are in the same page. On accessing {\tt x}, its page is swapped into the page buffer. On accessing {\tt y} and {\tt z}, the task need only access the volatile page buffer. The task updates the page, leaving it dirty at the task's completion. When the task completes, \sys first copies the page from the page buffer into the non-volatile {\tt pagesTemp} region before commit. Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.
